---
title: "Tooth Growth Analysis"
output: html_document
author: Chris Grafer
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(knitr)
library(ggplot2)
```


## Overview
The purpose of this paper is to analyze the ToothGrowth dataset in R and reach a conclusion on whether the data supports either the claim that the delivery method and/or dosage affect tooth growth in guinea pigs.

## Data Summary
The ToothGrowth data set contains data from an experiment where 60 guinea pigs were divided into 10 groups and provided with doses of either Vitaman C (VC) or Orange Juice (OJ) in three different doses (0.5, 1.0, 2.0 mg / day). The tooth length was then measured. The following scatterplot shows the length by supplement and dosage:

```{r echo=FALSE}
qplot(x = dose,  
      y = len, 
      data = ToothGrowth, 
      facets = . ~ supp, 
      xlab = "Dosage (mg/day)", 
      ylab = "Length of Tooth Growth", 
      colour = supp)
```

The following table summarizes the relevant statistics:


```{r echo = FALSE}
smry <- summarize(group_by(ToothGrowth, supp, dose), 
                  m = mean(len), 
                  sd = sqrt(var(len)), 
                  min = min(len), 
                  max = max(len))
col_names <- c("Supplement", "Dosage",
               "Mean", "Standard Deviation",
               "Minimum", "Maximum")
kable(smry, digits = 3, col.names = col_names)
```


## Tooth Growth Related to Supplement
To evaluate whether the data supports the claim that the type of supplement influences tooth growth, we will use the t-statistic to construct confidence intervals. We will assume the the groups that recieve Vitaman C are independent of the groups that recieve orange juice, and that each group has it's own variance.  The relevant t-test is calculated as:

```{r}
supp_test <- t.test(len ~ supp, data= ToothGrowth, paired = FALSE, var.equal = FALSE)
```
Since we are performing a two group test, the 95% confidence interval will tell us whether the difference in the means is statistically significant at the 5% level. Based on the fact that zero is within the 95% confidence interval, the differnece in means is not statistically signficant. 
```{r echo = FALSE}

supp_names <- c("Mean OJ", "Mean VC", "Difference", "95% Confidence Interval")
x<-  cbind(supp_test$estimate[1],
            supp_test$estimate[2],
            supp_test$estimate[1] - supp_test$estimate[2])
x <- format(x, digits = 6)
x <- cbind(x, paste(format(supp_test$conf[1], digits = 6),
                " to ",
                format(supp_test$conf[2], digits = 6)))
rownames(x) <- NULL
kable(x, digits = 3, col.names = supp_names)

```


## Tooth Growth Related to Dosage
We can perform t-tests and compare the various dosage levels and determine whether the result is statistaically significant at the 95% confidence interval.
```{r echo = FALSE}
low_dose <- ToothGrowth %>% filter(dose %in% c(1.0, 0.5))
high_dose <- ToothGrowth %>% filter(dose %in% c(2.0, 1.0))
ext_dose <- ToothGrowth %>% filter(dose %in% c(2.0, 0.5))
supp_test1 <- t.test(len ~ dose, data= low_dose, paired = FALSE, var.equal = FALSE)
supp_test2 <- t.test(len ~ dose, data= high_dose, paired = FALSE, var.equal = FALSE)
supp_test3 <- t.test(len ~ dose, data= ext_dose, paired = FALSE, var.equal = FALSE)

print(supp_test1)
print(supp_test2)
print(supp_test3)

```




## Conclusion